Domain Specific Language for high performance computing

### 2024-02-13 paper/paper.tex
can do test

\documentclass[ams, a4j]{U-AizuGT}
\usepackage{pifont}
% \usepackage{graphicx}
\usepackage{cite}
\usepackage{mathtools}
% \usepackage{listings,jvlisting}
% \documentclass[a4j]{jarticle} %ここは関係ない
\usepackage{listings,jvlisting} 
\usepackage[dvipdfmx]{graphicx}
\usepackage{physics}

\usepackage[utf8]{inputenc}
\usepackage{booktabs} % きれいな表を作成するためのパッケージ

\bibliographystyle{ieicetr}

@@

% \usepackage{listings,jlisting}

% \lstset{%
%   language={C},
%   basicstyle={\small},%
%   identifierstyle={\small},%
%   commentstyle={\small\itshape},%
%   keywordstyle={\small\bfseries},%
%   ndkeywordstyle={\small},%
%   stringstyle={\small\ttfamily},
%   frame={tb},
%   breaklines=true,
%   columns=[l]{fullflexible},%
%   numbers=left,%
%   xrightmargin=0zw,%
%   xleftmargin=3zw,%
%   numberstyle={\scriptsize},%
%   stepnumber=1,
%   numbersep=1zw,%
%   lineskip=-0.5ex%
% }
\lstset{
  basicstyle={\ttfamily},
  identifierstyle={\small},
  commentstyle={\smallitshape},
  keywordstyle={\small\bfseries},
  ndkeywordstyle={\small},
  stringstyle={\small\ttfamily},
  frame={tb},
  breaklines=true,
  columns=[l]{fullflexible},
  numbers=left,
  xrightmargin=0zw,
  xleftmargin=3zw,
  numberstyle={\scriptsize},
  stepnumber=1,
  numbersep=1zw,
  lineskip=-0.5ex
}

@@

 With the definition of this variable and a description of the
mathematical equation using the variable, a code to calculate the
interparticle interaction can be generated.
これらをもとにPykerの言語仕様を説明する。

\subsection{language specification}
Pykerでは、大きく、変数定義と相互作用記述に分かれる。
まず変数定義について説明する。

 先ほどのセクションよりPykerの変数定義を以下に正規表現で示す。
% \begin{lstlisting}  
\["(EPI|EPJ|FORCE)?\:\text{ } (vec3<)?(F64)>?\:\text{ } \text{\textbackslash w+}"\]
% \end{lstlisting}
EPI,EPJ,FORCEはそれぞれ粒子i,粒子j,結果を保存する変数であることを示す。それ以外の変数は何も書かない
ベクトルはvec3で表す。
もしベクトルであれば'<>'の中に型を定義する。
型は倍精度小数だけでF64で示す。
最後に変数名を記述する。

次に相互作用記述の説明をする。
相互作用の記述に使える演算は以下の図に示す。

以下に\label{eq:gravity}の計算をPyker記述した例を示す。
\begin{lstlisting}[frame=single, caption=gravity\_interparticle, label=fuga]
  EPI vec3<F64> ri
  EPJ vec3<F64> rj
  EPJ F64 mass
  EPJ F64 eps2
  FORCE vec3<F64> ai
  rij = ri - rj
  r2 = rij * rij + eps2
  r_inv = 1.0 / sqrt(r2)
  r2_inv = r_inv * r_inv
  mr_inv = mass * r_inv
  mr3_inv = r2_inv * mr_inv
  ai = mr3_inv * rij
\end{lstlisting}

5行目までは変数定義部になる。

\begin{table}[ht]
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  項目 & 値 & 単位 \\
  \midrule
  項目1 & 100 & kg \\
  項目2 & 200 & g \\
  項目3 & 300 & mg \\
\bottomrule
\end{tabular}
\end{table}

\section{Construction}
Pykerでは、どういうことができるのか？

We use Sympy, a library in Python for implementing  the implementation of pyker.
SymPy supports various operations, and the formulas using these operations are internally treated as 
syntax trees. We leveraged this functionality to read formulas and generate code capable of parallel execution.
For parallelization, we used an instruction set named Advanced Vector Extensions 2 \lparen AVX2\rparen. AVX2 is one of the 
extended instruction sets implemented in Intel's CPUs. The next section will explain how parallelization is carried out using AVX2.

@@

\begin{tabular}{|l|r|r|} \hline
  N & Pyker & PIKG \\ \hline
  25000 & 6.725156 & 4.444766 \\
  10000 & 0.738918sec & 0.517324sec \\

 \\ \hline
\end{tabular}

@@

@@

 \\ \hline
\end{tabular}

>>>>

\documentclass[ams, a4j]{U-AizuGT}
\usepackage{pifont}
% \usepackage{graphicx}
\usepackage{cite}
\usepackage{mathtools}
% \usepackage{listings,jvlisting}
% \documentclass[a4j]{jarticle} %ここは関係ない
\usepackage{listings,jvlisting} 
\usepackage[dvipdfmx]{graphicx}
\usepackage{physics}
\usepackage{tikz}
\usepackage[utf8]{inputenc}
\usepackage{booktabs} % きれいな表を作成するためのパッケージ
\usetikzlibrary{shapes,arrows,positioning}
\bibliographystyle{ieicetr}
\usetikzlibrary{positioning, arrows.meta, shapes.geometric}

@@

\lstset{
  basicstyle={\ttfamily},
  identifierstyle={\small},
  commentstyle={\smallitshape},
  keywordstyle={\small\bfseries},
  ndkeywordstyle={\small},
  stringstyle={\small\ttfamily},
  frame={tb},
  breaklines=true,
  columns=[l]{fullflexible},
  numbers=left,
  xrightmargin=0zw,
  xleftmargin=3zw,
  numberstyle={\scriptsize},
  stepnumber=1,
  numbersep=1zw,
  lineskip=-0.5ex
}

@@

 With the definition of this variable and a description of the
mathematical equation using the variable, a code to calculate the
interparticle interaction can be generated.
We explain the language specification in the next section.

\subsection{language specification}
There are two parts in Pyker. variable definition part and 
interaction description part.
First, we explain about the variable definition part.

\subsubsection{variable definition part}
Variable definitions are shown below in regular expressions.
% \begin{lstlisting}  
\["(EPI|EPJ|FORCE)?\:\text{ } (vec3<)?(F64)>?\:\text{ } \text{\textbackslash w+}"\]
% \end{lstlisting}
EPI and EPJ represent the variables for particles i and j, 
respectively, and FORCE indicates that they are variables that
 store the results.   Other variables are not written.

 Vector are denoted by vec3.If variable is a vector, the variable type
 is defined in '<>'.

Variable type is 'F64' for double-precision.In Pyker, Variable type is only 'F64'.

Finally, the name of the variable is written.

次に相互作用記述の説明をする。
相互作用の記述に使える演算は以下の図に示す。
\begin{lstlisting}[frame=single]
 +, -, /, *, **, sqrt(), =, +=
\end{lstlisting}
$+, -, *, /$ はそれぞれ 加算、減算、乗算、徐算を表す。
$**$ はべき乗を表し、sqrt()は平方根を表す。
$*$の項がどちらもベクトルの時、$**$の底がベクトル、指数が2の場合にそれぞれ内積の計算と判断される。
$=$は一時変数を定義する際に使用する。一時変数は定義する際に型を定義する必要はなく、すでに定義されている変数の型とベクトルから
推論を行い決定する。
$+=$はFORCEの変数に計算結果を加算することを示す。
原則、要素がベクトルの場合は、ベクトルの要素ごとに演算を行う。
記述に使えるオブジェクトは、変数定義を行った変数とその記述より前に定義された一時変数と浮動小数である。

以下に\label{eq:gravity}の計算をPyker記述した例を示す。
\begin{lstlisting}[frame=single, caption=gravity\_interparticle, label=fuga]
  EPI vec3<F64> ri
  EPJ vec3<F64> rj
  EPJ F64 mass
  EPJ F64 eps2
  FORCE vec3<F64> ai
  rij = ri - rj
  r2 = rij * rij + eps2
  r_inv = 1.0 / sqrt(r2)
  r2_inv = r_inv * r_inv
  mr_inv = mass * r_inv
  mr3_inv = r2_inv * mr_inv
  ai += mr3_inv * rij
\end{lstlisting}

次に、これらの記述からSIMD命令を行えるコードを生成する説明をする。

\section{Parser of Pyker}
Pykerでは、
このセクションでは、まずどのようにしてSIMDによる並列化を行うか説明し、その後Pykerの記述の読み込みからコード生成までの処理内容と流れを説明します。

\begin{tikzpicture}
  \tikzset{
    block/.style={
      rectangle, draw, text centered, text width=3cm,
      minimum height=1.5cm
    },
    line/.style={
      draw, very thick, color=black!50, -Latex
    }
  }

  % Place nodes
  \node[block] (parse) {Parse};
  \node[block, right=of parse] (type_inference) {type\_inference};
  \node[block, right=of type_inference] (bitree) {bitree};
  \node[block, right=of bitree] (codegen) {Codegen};

  % Draw edges
  \path[line] (parse) -- (type_inference);
  \path[line] (type_inference) -- (bitree);
  \path[line] (bitree) -- (codegen);
\end{tikzpicture}

\subsubsection{AVX2 for parallelization}
% _mm256dの説明はここですべきか？ codegen で説明すべきか
To perform SIMD operations, just like with regular computations, data is first loaded from memory, computed based on that data,
and the results are stored. However, in the case of SIMD, computations are performed using SIMD-specific registers, so the 
methods of loading, computing, and storing differ. When loading, multiple elements that will undergo the same instruction are
stored in a SIMD-specific register. These are then used to perform computations simultaneously, and the results are stored for
each.

For instance, consider a calculation like C = X + Y, where elements of double-precision floating arrays x and y are computed
. When using AVX2 for SIMD, since the registers are 256 bits, four elements from x and y are loaded each. These are then added 
simultaneously, and the results are stored in the array C. When loading into SIMD registers, if it's a 1-dimensional array, 
contiguous memory access is sufficient, but for data structures like Array of Structures \lparen AoS \rparen shown in Figure a., elements need
to be loaded individually.

For example, when considering the 3-dimensional coordinates of particles as double-precision floats 
in pos\lbrack n\rbrack \lbrack 3\rbrack  \lparen where n is the total number of particles, and pos\lbrack i \rbrack \lbrack 0\rbrack , pos\lbrack i\rbrack \lbrack 1\rbrack , 
pos\lbrack i\rbrack \lbrack 2\rbrack  represent the x, y, z coordinates 
of the i-th particle, respectively\rparen, and calculating the difference in x coordinates between particles i and j, it becomes dx = 
pos\lbrack i\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack . However, for parallelization with SIMD, desiring dx0 = pos\lbrack i\rbrack \lbrack 0\rbrack  - pos\lbrack j \rbrack \lbrack 0\rbrack , dx1 = pos\lbrack i+1\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack 
, dx2 = pos\lbrack i+2\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack , dx3 = pos\lbrack i+3\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack , 
the data are packed as \lparen pos\lbrack i\rbrack \lbrack 0\rbrack , pos\lbrack i+1\rbrack \lbrack 0\rbrack , pos\lbrack i+2\rbrack \lbrack  0\rbrack , pos\lbrack  i+3\rbrack \lbrack  0\rbrack \rparen,
\lparen pos\lbrack  j\rbrack \lbrack  0\rbrack , pos\lbrack  j\rbrack \lbrack  0\rbrack , pos\lbrack  j\rbrack \lbrack  0\rbrack , pos\lbrack  j\rbrack \lbrack  0\rparen, \lparen dx0, dx1, dx2, dx3\rparen and subtraction is performed simultaneously. However, since data needs
to be loaded onto SIMD registers all at once, the gather instruction is used to load data for particle i. The gather instruction is used
to read non-contiguous data elements from memory addresses by specifying addresses. An explanation of the gather instruction is shown in 
Figure x. Data loaded in this way are computed simultaneously using SIMD instructions.

\begin{lstlisting}[frame=single]
  dx0 = pos[i][0] - pos[j][0]
  dx1 = pos[i + 1][0] - pos[j][0]
  dx2 = pos[i + 2][0] - pos[j][0]
  dx3 = pos[i + 3][0] - pos[j][0]
\end{lstlisting}

\begin{lstlisting}[frame=single]
  (x1, x2, x3, x4) + (y1, y2, y3, y)

When storing computation results, addresses may 
not be contiguous. In such cases, there is an instruction called scatter, which is the counterpart to gather, but it is not supported 
in AVX2, so data are stored using pointers individually. In this way, parallelization is achieved with SIMD. Moreover, these instructions
can be explicitly handled in high-level programming languages like C and C++ using intrinsic functions. To handle intrinsic functions 
in C and C++, include the file immintrin.h, which is available as a standard in the language. An example of C++ code that calculates the difference in x coordinates, simd\_tmp.cpp, is shown. { Pyker}  generates C++ code capable of executing SIMD instructions like simd\_tmp.cpp
from the described formulas. The next section will explain the Pyker.

@@

\begin{tabular}{|l|r|r|} \hline
  N & Pyker & PIKG \\ \hline
  25000 & 6.725156 & 4.444766 \\
  10000 & 0.738918sec & 0.517324sec \\
 \\ \hline
\end{tabular}

@@

\subsection{Hermite}

@@

\begin{table}
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  N & Pyker & PIKG \\
  \midrule
  50000 & 5.282461100 & 4.059262 \\
  25000 & 1.322715200 & 1.008046 \\
  10000 & 0.213696 & 0.161884 \\
  1000 & 0.002116 & 0.001713 \\
\bottomrule
\end{tabular}
\end{table}
\begin{table}[ht]
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  項目 & 値 & 単位 \\
  \midrule
  項目1 & 100 & kg \\
  項目2 & 200 & g \\
  項目3 & 300 & mg \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  N & Pyker & PIKG \\
  \midrule
  50000 & 24.755255 & 22.767920 \\
  25000 & 6.136025 & 5.740028 \\
  10000 & 1.044833 & 0.901585\\
  1000 & 0.010176 & 0.009109 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  N & Pyker & PIKG \\
  \midrule
  50000 & 13.796789& 10.794405\\
  25000 & 3.472925 & 2.693424\\
  10000 & 0.557972 & 0.424820\\
  1000 & 0.005405& 0.004240\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  N & Pyker & PIKG \\
  \midrule
  50000 & 2.476695 & 2.439798\\
  25000 & 0.617220 & 0.598798\\
  10000 & 0.094859 & 0.094860\\
  1000 & 0.000939& 0.000988\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  N & Pyker & PIKG \\
  \midrule
  50000 & 2.153937 & 2.194700\\
  25000 &0.542478 & 0.554280\\
  10000 & 0.085537 &0.087284\\
  1000 & 0.000869&0.000874\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  N & Pyker & PIKG \\
  \midrule
  50000 & 15.103493& 9.904044\\
  25000 &3.736596 & 2.480025\\
  10000 &0.543367 &0.398513\\
  1000 & 0.005392&0.003911\\
\bottomrule
\end{tabular}
\end{table}
